---
title: "Democracy and Power Diffusion"
author: "Julian Bernauer"
date: "July 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How to cite 

Bernauer, Julian and Adrian Vatter (Forthcoming). Power Diffusion and Democracy: Institutions, Deliberation and Outcomes. Cambridge: Cambridge University Press. 


## Factor Analytical Item Response Theoretical (FAIRT) Approach 
Bernauer and Vatter (2018) propose an approach which 
1) does more justice to the measurement levels of the indicators used,
2) let's their discriminatory power vary,
2) inherently estimates scores 
3) considers uncertainty

This involves a healthy junk of code, which involves quite some repetition, though. Just scroll further down if you're not interested! 

```{r fairt}
model <- "model{
#Election-level model for electoral disproportionality and the number of parties
for(k in 1:NE){
elec[k] ~ dnorm(mu.elec[k],tau.elec)
mu.elec[k] <- alpha.elec[idne[k]] + gamma.elec1*ex1[k]
par[k] ~ dnorm(mu.par[k],tau.par)
mu.par[k] <- alpha.par[idne[k]] + gamma.par1*ex1[k]
cab[k] ~ dcat(p.cab[k,1:4])
p.cab[k,1] <- Q[k,1]
p.cab[k,2] <- Q[k,2] - Q[k,1]
p.cab[k,3] <- Q[k,3] - Q[k,2]
p.cab[k,4] <- 1 - Q[k,3]
for (k.cut in 1:3){
logit(Q[k,k.cut]) <- C[idne[k],k.cut] - beta.cab1*ex1[k]
}
}
#Country-level model for electoral disproportionality and the number of parties
for(j in 1:N){
alpha.elec[j] ~ dnorm(mu.ae[j],tau.ae)
mu.ae[j] <- ae + a.elec1*cx1[j]
alpha.par[j] ~ dnorm(mu.ap[j],tau.ap)
mu.ap[j] <- ap + a.par1*cx1[j]
}
ae ~ dnorm(0,.0001)
a.elec1 ~ dnorm(0,.0001)
tau.ae <- pow(sigma.ae, -2)
sigma.ae ~ dunif(0, 50)
ap ~ dnorm(0,.0001)
a.par1 ~ dnorm(0,.0001) I(0,)
tau.ap <- pow(sigma.ap, -2)
sigma.ap ~ dunif(0, 50)
tau.elec <- pow(sigma.elec, -2)
sigma.elec ~ dunif(0, 50)
tau.par <- pow(sigma.par, -2)
sigma.par ~ dunif(0, 50)
gamma.elec1 ~ dnorm(0, .001)
gamma.par1 ~ dnorm(0, .001) I(0,)
#Election- and country level ordinal logit model for cabinet type
for (j in 1:N){
C[j,1] ~ dnorm (mu.ac[j], tau.ac)
mu.ac[j] <- acab - (a.cab1*cx1[j] + beta.prescab*pres[j] + beta.dircab*dir[j])
delta[j,1] ~ dexp(2)
delta[j,2] ~ dexp(2)
C[j, 2] <- C[j,1] + delta[j,1]
C[j, 3] <- C[j,2] + delta[j,2]
}
acab ~ dnorm(0,.0001)
tau.ac <- pow(sigma.ac,-2)
sigma.ac ~ dunif(0,50)
a.cab1 ~ dnorm(0, .001)
beta.cab1 ~ dnorm(0, .001)
#Country-year- and country-level model for fiscal decentralization
for(i in 1:NT){
dec[i] ~ dnorm(mu.dec[i],tau.dec)
mu.dec[i] <- alpha.dec[idn[i]] + gamma.dec2*x2[i]
}
for(j in 1:N){
alpha.dec[j] ~ dnorm(mu.ad[j],tau.ad)
mu.ad[j] <- adec + a.dec2*cx2[j]
}
adec ~ dnorm(0,.0001)
a.dec2 ~ dnorm(0,.0001)
tau.ad <- pow(sigma.ad, -2)
sigma.ad ~ dunif(0, 50)
tau.dec <- pow(sigma.dec, -2)
sigma.dec ~ dunif(0, 50)
gamma.dec2 ~ dnorm(0, .001) I(0,)
#Country-level models for the remaining indicators
for(i in 1:N){
exeleg[i] ~ dnorm(mu.exeleg[i],tau.exeleg)
mu.exeleg[i] <- alpha.exeleg + gamma.exeleg1*cx1[i] + gamma.presexeleg*pres[i]
dir[i] ~ dnorm(mu.dir[i],tau.dir)
mu.dir[i] <- alpha.dir + gamma.dir4*cx4[i]
pres[i] ~ dcat(p.pres[i,1:3])
mu.pres[i] <- beta.pres3*cx3[i]
logit(Q.pres[i,1]) <- tau.pres[1] - mu.pres[i]
p.pres[i,1] <- Q.pres[i,1]
logit(Q.pres[i,2]) <- tau.pres[2] - mu.pres[i]
p.pres[i,2] <- Q.pres[i,2] - Q.pres[i,1]
p.pres[i,3] <- 1 - Q.pres[i,2]
fed[i] ~ dcat(p.fed[i,1:3])
mu.fed[i] <- beta.fed2*cx2[i]
logit(Q.fed[i,1]) <- tau.fed[1] - mu.fed[i]
p.fed[i,1] <- Q.fed[i,1]
logit(Q.fed[i,2]) <- tau.fed[2] - mu.fed[i]
p.fed[i,2] <- Q.fed[i,2] - Q.fed[i,1]
p.fed[i,3] <- 1 - Q.fed[i,2]
bic[i] ~ dcat(p.bic[i,1:4])
mu.bic[i] <- beta.bic2*cx2[i]
logit(Q.bic[i,1]) <- tau.bic[1]-mu.bic[i]
p.bic[i,1] <- Q.bic[i,1]
logit(Q.bic[i,2]) <- tau.bic[2] - mu.bic[i]
p.bic[i,2] <- Q.bic[i,2] - Q.bic[i,1]
logit(Q.bic[i,3]) <- tau.bic[3] - mu.bic[i]
p.bic[i,3] <- Q.bic[i,3] - Q.bic[i,2]
p.bic[i,4] <- 1 - Q.bic[i,3]
const[i] ~ dcat(p.const[i,1:3])
mu.const[i] <- beta.const2*cx2[i]
logit(Q.const[i,1]) <- tau.const[1] - mu.const[i]
p.const[i,1] <- Q.const[i,1]
logit(Q.const[i,2]) <- tau.const[2] - mu.const[i]
p.const[i,2] <- Q.const[i,2] - Q.const[i,1]
p.const[i,3] <- 1 - Q.const[i,2]
jud[i] ~ dcat(p.jud[i,1:3])
mu.jud[i] <- beta.jud2*cx2[i]
logit(Q.jud[i,1]) <- tau.jud[1] - mu.jud[i]
p.jud[i,1] <- Q.jud[i,1]
logit(Q.jud[i,2]) <- tau.jud[2] - mu.jud[i]
p.jud[i,2] <- Q.jud[i,2] - Q.jud[i,1]
p.jud[i,3] <- 1 - Q.jud[i,2]
}
tau.exeleg <- pow(sigma.exeleg, -2)
sigma.exeleg ~ dunif(0, 50)
tau.dir <- pow(sigma.dir, -2)
sigma.dir ~ dunif(0, 50)
alpha.exeleg ~ dnorm(0, .001)
alpha.dir ~ dnorm(0, .001)
gamma.exeleg1 ~ dnorm(0, .001)
gamma.dir4 ~ dnorm(0, .001)
tau.pres[1] ~ dnorm(0,.01)
for (j in 1:1){
delta.pres[j] ~ dexp(2)
tau.pres[j+1] <- tau.pres[j] + delta.pres[j]
}
tau.fed[1] ~ dnorm(0,.01)
for (j in 1:1){
delta.fed[j] ~ dexp(2)
tau.fed[j+1] <- tau.fed[j] + delta.fed[j]
}
tau.bic[1] ~ dnorm(0,.01)
for (j in 1:2){
delta.bic[j] ~ dexp(2)
tau.bic[j+1] <- tau.bic[j] + delta.bic[j]
}
tau.jud[1] ~ dnorm(0,.01)
for (j in 1:1){
delta.jud[j] ~ dexp(2)
tau.jud[j+1] <- tau.jud[j] + delta.jud[j]
}
tau.const[1] ~ dnorm(0,.01)
for (j in 1:1){
delta.const[j] ~ dexp(2)
tau.const[j+1] <- tau.const[j] + delta.const[j]
}
beta.fed2 ~ dnorm(0, .001)
beta.bic2 ~ dnorm(0, .001) I(0,)
beta.const2 ~ dnorm(0, .001)
beta.jud2 ~ dnorm(0, .001)
beta.pres3 ~ dnorm(0, .001)
#Priors for the country scores of power diffusion at the appropriate levels of
#elections (proportional power diffusion),
for(k in 1:NE){
ex1[k] ~ dnorm(cx1[idne[k]],1)
}
#country-years (decentral power diffusion),
for(i in 1:NT){
x2[i] ~ dnorm(cx2[idn[i]],1)
}
#or countries (all variants)
for(k in 1:55){
cx1[k] ~ dnorm(0,1)
cx2[k] ~ dnorm(0,1)
cx3[k] ~ dnorm(0,1)
cx4[k] ~ dnorm(0,1)
}
#Restriction of country-level proportional power diffusion for Switzerland
cx1[56] ~ dnorm(0,1) I(0,)
cx1[57] ~ dnorm(0,1)
cx1[58] ~ dnorm(0,1)
cx1[59] ~ dnorm(0,1)
cx1[60] ~ dnorm(0,1)
cx1[61] ~ dnorm(0,1)
#Restriction of country-level decentral power diffusion for Switzerland
cx2[56] ~ dnorm(0,1) I(0,)
cx2[57] ~ dnorm(0,1)
cx2[58] ~ dnorm(0,1)
cx2[59] ~ dnorm(0,1)
cx2[60] ~ dnorm(0,1)
cx2[61] ~ dnorm(0,1)
cx3[56] ~ dnorm(0,1)
cx3[57] ~ dnorm(0,1)
cx3[58] ~ dnorm(0,1)
cx3[59] ~ dnorm(0,1)
#Restriction of country-level presidential power diffusion for the United States
cx3[60] ~ dnorm(0,1) I(0,)
cx3[61] ~ dnorm(0,1)
#Restriction of country-level direct power diffusion for Switzerland
cx4[56] ~ dnorm(0,1) I(0,)
cx4[57] ~ dnorm(0,1)
cx4[58] ~ dnorm(0,1)
cx4[59] ~ dnorm(0,1)
cx4[60] ~ dnorm(0,1)
cx4[61] ~ dnorm(0,1)
#Collected cross-indicator effects
beta.dircab ~ dnorm(0, .001)
beta.prescab ~ dnorm(0, .001)
#Restriction of relationship between presidentialism and parliamentary power
gamma.presexeleg ~ dnorm(0, .001) I(,0)
}"
```
